# Data

## Sources

Our original data file, `Hate_Crime.csv`, is from the FBI's Crime Data Explorer (CDE) (<https://crime-data-explorer.fr.cloud.gov/pages/home>). According to the description by the FBI, the data was collected through the FBI's Uniform Crime Reporting (UCR) Program. UCR is a program aimed at providing reliable crime statistics for public use. Approximately 18,000 law enforcement units in the US participate in this program. Each unit voluntarily sends its data through the National Incident-Based Reporting System (NIBRS), which allows them to provide detailed information about a crime, such as offenders' bias against specific social identity. At the end of each year, the FBI will release the data a year before with multiple sections through CDE, and Hate Crime Statistics is a dedicated collection along with others.

```{r}
library(tidyverse)
hate = read.csv("resource/hate_crime.csv")
str <- sprintf("The dimension of the dataset is %d rows and %d columns.\n",
               dim(hate)[1], dim(hate)[2])
cat(str)
cat("The columns' names are:\n")
names(hate)
```

As we discover, the data has 219577 rows and 28 columns. Most of the variables are categorical. `INCIDENT_ID`, `ADULT_VICTIM_COUNT`, `JUVENILE_VICTIM_COUNT`, `TOTAL_OFFENDER_COUNT`, `ADULT_OFFENDER_COUNT`, `JUVENILE_OFFENDER_COUNT`, `VICTIM_COUNT`, and `TOTAL_INDIVIDUAL_VICTIMS` are continuous variables. `DATA_YEAR` and `INCIDENT_DATE` should be date variables, but they are stored in `int` and `chr`.

Although the FBI provided an extensive and detailed dataset, it contains several issues. The most important one, as stated by the FBI, is that because a crime's motivation is subjective, it is difficult to determine whether the crime is because of the offenders' bias. Hence, law enforcement concludes a hate crime only if sufficient evidence proves it is a hate crime. This results in underestimating the overall occurrence of hate crimes, but it could guarantee the accuracy of our analysis. As we understood and interpreted the dataset, we found some problems regarding the quality of the dataset. Multiple columns contain value of empty string instead of `NA`, which leads to erroneous missing value analysis. Some observation contains multiple variables in the `OFFENSE_NAME` and `BIAS_DESC` columns. These situations are indicated in `MULTIPLE_OFFENSE` and `MULTIPLE_BIAS`. It is reasonable for a crime aimed at multiple groups of people with several offenses, but it creates obstacles when analyzing offense type and bias. Other problems, like `TOTAL_OFFENDER_COUNT` contains a massive portion of 0s, and adults and juveniles partition is not reported before 2013. Since we dropped these columns, we did not dig into the reason for these problems.

## Cleaning / transformation

The first cleaning step is to extract the columns needed for our target. We have already mentioned our goal in the introduction section. Here we selected related columns and reduced it to 7 columns.

```{r}
hate_select <- hate |>
  select(c(INCIDENT_ID, STATE_NAME, REGION_NAME, INCIDENT_DATE,
           OFFENDER_RACE, OFFENSE_NAME, BIAS_DESC))
hate_clean_na <- hate_select
str <- sprintf("The dimension of the dataset is %d rows and %d columns.\n", 
        dim(hate_select)[1], dim(hate_select)[2])
cat(str)
cat("The columns' names are:\n")
names(hate_select)
```

Next, we will split the row with multiple `BIAS_DESC` and `OFFENSE_NAME`. Before the split, the unique entries in `BIAS_DESC` and `OFFENSE_NAME` are 279 and 353. After the split, the unique entries in `BIAS_DESC` and `OFFENSE_NAME` are reduced to 35 and 48. Note that this will create multiple rows for one crime and can lead to overcounting. Therefore, we will only count crimes with distinct `INCIDENT_ID` to overcome this issue.

```{r}
cat("BEFORE\n")
str <- sprintf("The unique BIAS_DESC counts is: %d.\n",
               length(unique(hate_select$BIAS_DESC)))
cat(str)
str <- sprintf("The unique OFFENSE_NAME counts is: %d.\n",
        length(unique(hate_select$OFFENSE_NAME)))
cat(str)
hate_split <- hate_select |>
  separate_rows(BIAS_DESC, sep = ";") |>
  separate_rows(OFFENSE_NAME, sep = ";")
cat("AFTER\n")
str <- sprintf("The unique BIAS_DESC counts is: %d.\n",
        length(unique(hate_split$BIAS_DESC)))
cat(str)
str <- sprintf("The unique OFFENSE_NAME counts is: %d.\n",
        length(unique(hate_split$OFFENSE_NAME)))
cat(str)
str <- sprintf("The dimension of the dataset is %d rows and %d columns.\n", 
        dim(hate_split)[1], dim(hate_split)[2])
cat(str)
```

A strange pattern in `BIAS_DESC` is that while it contains `Anti-Lesbian (Female)`, `Anti-Gay (Male)`, `Anti-Bisexual`, and `Anti-Transgender` separately, it also has an entry `Anti-Lesbian, Gay, Bisexual, or Transgender (Mixed Group)`. We will combine the separate listings into the mixed one.

```{r}
hate_comb <- hate_split |>
  mutate(BIAS_DESC =
           replace(BIAS_DESC, 
                   BIAS_DESC == "Anti-Lesbian (Female)",
                   "Anti-Lesbian, Gay, Bisexual, or Transgender (Mixed Group)")) |>
  mutate(BIAS_DESC =
           replace(BIAS_DESC, 
                   BIAS_DESC == "Anti-Gay (Male)",
                   "Anti-Lesbian, Gay, Bisexual, or Transgender (Mixed Group)")) |>
  mutate(BIAS_DESC =
           replace(BIAS_DESC, 
                   BIAS_DESC == "Anti-Bisexual",
                   "Anti-Lesbian, Gay, Bisexual, or Transgender (Mixed Group)")) |>
  mutate(BIAS_DESC =
           replace(BIAS_DESC, 
                   BIAS_DESC == "Anti-Transgender",
                   "Anti-Lesbian, Gay, Bisexual, or Transgender (Mixed Group)"))
str <- sprintf("The unique BIAS_DESC counts is: %d.\n",
               length(unique(hate_comb$BIAS_DESC)))
cat(str)
```

The `INCIDENT_DATE` is stored as `character` class. We need to change it into `Date` class.

```{r}
cat("BEFORE\n")
str <- sprintf("The class of INCIDENT_DATE is : %s.\n",
               class(hate_comb$INCIDENT_DATE))
cat(str)
hate_clean <- hate_comb |>
  mutate(INCIDENT_DATE = as.Date(INCIDENT_DATE, "%d-%b-%y"))
cat("AFTER\n")
str <- sprintf("The class of INCIDENT_DATE is : %s.\n",
               class(hate_clean$INCIDENT_DATE))
cat(str)
```

The last step is to process empty strings. We consider it as `NA`. And because empty string only happens in categorical variables, we will replace it with `Unknown`.

```{r}
cat("BEFORE\n")
str <- sprintf("The number of rows with empty string is : %d.\n",
               dim(filter_all(hate_clean, any_vars(. == "")))[1])
cat(str)
hate_clean[hate_clean == ""] <- "Unknown"
cat("AFTER\n")
str <- sprintf("The number of rows with empty string is : %d.\n",
               dim(filter_all(hate_clean, any_vars(. == "")))[1])
cat(str)
```

## Missing value analysis

Since our data set is gigantic, we cannot use either a heat map or `mi` library to plot and interpret the missing pattern. Here we will use `redav` library provided by Professor Robbins to plot the missing patterns. We changed the column names into their index to better visualize the graph. We also transform all empty strings into `NA`.

```{r}
library(redav)
hate_na <- hate
colnames(hate_na) <- as.character(c(1:28))
hate_clean_na <- hate_na |>
  select(c(1, 8, 10, 13, 19, 22, 25))
hate_na[hate_na == ""] <- NA
hate_clean_na[hate_clean_na == ""] <- NA
```

For the raw dataset, the missing pattern seems messy. However, it shows that about 95% of rows are missing `PUB_AGENCY_UNIT`, 80% of rows are missing `OFFENDER_ETHNICITY`, and 75% are missing `ADULT_VICTIM_COUNT`, `JUVENILE_VICTIM_COUNT`, `ADULT_OFFENDER_COUNT`, and `JUVENILE_OFFENDER_COUNT`. It is not surprising that `PUB_AGENCY_UNIT` has a considerable number of missing because after we dig into the dataset, `PUB_AGENCY_UNIT` are only provided when the unit is affiliated with home law enforcement, like university or state police. Aside from that, the other five columns are usually missing together. While checking the dataset, we found that all the rows are missing these five values until 2013. After 2013, some of them are still missing but gradually decreased. A possible explanation is that before 2013, the FBI did not require law enforcement to report this data. After 2013, the FBI required law enforcement to provide this information, and some of them responded to it immediately, while others may have had some difficulties in submitting this data. Still, eventually, they all comply with the FBI requirement. In addition to these major columns missing, `TOTAL_INDIVIDUAL_VICTIMS` and `OFFENDER_RACE` also have some minor missing, but they don't have correlation, which seems to be happened completely at random.

```{r}
plot_missing(hate_na)
```

The cleaned data only has one missing pattern and a negligible amount of missing. The missing column is `OFFENDER_RACE` and in the cleaning part, we have already counted there only 20 missing values. We found that `OFFENDER_RACE` has an entry `Unknown`. Therefore, we will replace `NA` with `Unknown`.

```{r}
plot_missing(hate_clean_na)
```