# Results

## Trends of hate crime

The following shows the trend chart of the total number of reported hate crime cases with each year. As can be seen from the figure, hate crime case increased most sharply in 1991-1992, 1994-1995, 2000-2001 and 2019-2020. Conversely, from 1993-1994, 2001-2002, 2008-2009, 2012-2013, the decline was the most dramatic.

From 2001 to 2014, the total number of hate crime cases showed a general trend of decline. However, from 2014 to 2020, the total number of hate crime cases in the United States increased dramatically.

In the whole chart, 1991 had the lowest total number of hate crime cases. 2020 has the highest total number of hate crime cases. 

We know that the year has something to do with how long a U.S. president has been in office, Based on this picture, we will answer the question of which presidential administration had the highest or lowest hate crime incidents in the third subsection. 


```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggthemes)
library(plotly)
hate_plusyr = hate_clean %>% mutate(DATA_YEAR = year(INCIDENT_DATE))
hate_plusyr_1 = hate_plusyr[!duplicated(hate_plusyr[,c('INCIDENT_ID','DATA_YEAR')]),]
year <- hate_plusyr_1 %>% select(DATA_YEAR) %>%
 group_by(DATA_YEAR) %>%
 summarize(count = sum(n()))

#trends_plot = ggplot(year, aes(DATA_YEAR,count))+
#  geom_line()+
#  scale_x_continuous(breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020))+
#    labs(title="Total Hate Crime line chart for 1991-2020",x="Year", y = "Total case for reported Hate Crimes")+
#  theme(plot.title = element_text(hjust = 0.5))+ theme_economist() 
#trends_plot

trends_plot<- plot_ly(year, x = year$DATA_YEAR, y = year$count)
peak <- dplyr::filter(year, count == max(count))
troughs <- dplyr::filter(year, count == min(count))
second_peak = dplyr::filter(year, DATA_YEAR == 2001)
second_troughs = dplyr::filter(year, DATA_YEAR == 2014)
trends_plot %>%
  add_lines(x = year$DATA_YEAR, y = year$count) %>%
  layout(title = "Total Hate Crime Line Chart for 1991-2020",
         showlegend = FALSE,xaxis = list(title = 'Year'), 
         yaxis = list(title = 'Total case for reported Hate Crimes'))%>%
  layout(annotations = list(x = peak$DATA_YEAR, y = peak$count, text = "First Peak", showarrow = T)) %>%
  layout(annotations = list(x = troughs$DATA_YEAR, y = troughs$count, text = "First Trough", showarrow = T)) %>%
  layout(annotations = list(x = second_peak$DATA_YEAR, y = second_peak$count, text = "Second Peak", showarrow = T))%>%
  layout(annotations = list(x = second_troughs$DATA_YEAR, y = second_troughs$count, text = "Second Trough", showarrow = T))  


```


## Which groups of people were attacked the most?

The plot shows the number of total hate crimes for each prejudice. As can be seen from the Cleveland dot chart, black or African Americans are the groups most attacked by hate crime(almost 70,000 reported incidents). Asians, however, were the seventh most attacked group(over 7,000 reported incidents). Please note that our graph here only shows the total number of incidents involving people who have been attacked over the last 20 years, as opposed to which group is most vulnerable. We will take the results of the graph to analyse the relationship between the total number of people attacked in different categories under different presidents. 

```{r, fig.height = 10, fig.width = 8, fig.align = "center"}
bias <- hate_clean %>% select(BIAS_DESC) %>%
 group_by(BIAS_DESC) %>%
 summarize(count = sum(n()))%>%
 arrange(desc(count))

#ggplot(bias, aes(reorder(BIAS_DESC, (count)), count, fill = BIAS_DESC))+
# geom_bar(stat = 'identity')+
# xlab(label = 'Hate Crime')+
# theme(axis.text.x = element_blank(), axis.ticks = element_blank())+
# theme(legend.key.size = unit(x = 2, units = 'line'),
#       legend.text = element_text(size = 14),
#       legend.title = element_text(size = 15))

ggplot(bias, aes(x = count, y = fct_reorder(BIAS_DESC, count))) +
  geom_point(color = "blue") +
  labs(title="Cleveland Dot Plot of Total Hate Crimes per Prejudice", x = "Hate Crime cases", y = "Prejudice")+  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5))

```

## The relationship between the race of the attacker and the race of the attacked.

The following graph shows the relationship between the race of the attacker and the race of the attacked. Note we will ignore the unknown offender race in this case. We found that when Black or African american people were attacked, white people attacked them the most, by a significant margin. We also found that whites were the most aggressive toward all groups except toward Jewish. The majority of Anti-White hate crimes were committed by Black people and White people. 

```{r, fig.height = 10, fig.width = 18, fig.align = "center"}
bias_top_7 = bias %>% head(7)
attacker_attacked_df = hate_clean %>% filter(BIAS_DESC %in% bias_top_7$BIAS_DESC) %>% left_join(bias_top_7)

attacker_attacked_plot = ggplot(attacker_attacked_df, aes(x=fct_rev(fct_reorder(BIAS_DESC,count,.desc = TRUE)), fill = OFFENDER_RACE)) +
  geom_bar(position = "dodge") +
  ggtitle("Offender Race per Prejudice (1991-2020)") +
  coord_flip() +
  ylab("Hate Crime cases") + xlab("Prejudice")+
  theme(plot.title = element_text(hjust = 0.5)) +theme(legend.key.size = unit(x = 2, units = 'line'),
       legend.text = element_text(size =15),
       legend.title = element_text(size = 18),title = element_text(size = 30)) +theme(strip.text.x = element_text(size = 25))+
theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=18))+theme(legend.position = c(0.8, 0.3))+ theme(legend.background = element_rect(fill="lightblue",
                                  linetype="solid", 
                                  colour ="darkblue")) +guides(fill=guide_legend(title="Offender Race"))

attacker_attacked_plot
```






## Which presidential administration had the highest or lowest hate crime rate?

Because each U.S. president has a different number of years in office, it would be unfair to compare total hate crime cases here. So we divided the number of hate crime cases during each president's tenure by the number of years in office to get the hate crime rate for each president. We found that Donald Trump had the highest hate prime rate and George Bush(1st term) had the lowest. But the difference between the three presidents with the highest rates is not obvious.

```{r}
president_ds <- hate_plusyr %>% 
mutate(president = as.factor(case_when(DATA_YEAR < 1993 ~ 'George Bush(1st term)',
                                       DATA_YEAR >= 1993 & DATA_YEAR <= 2000 ~ 'Bill Clinton',
                                       DATA_YEAR >= 2001 & DATA_YEAR <= 2008 ~ 'George Bush(2nd term)',
                                       DATA_YEAR >= 2009 & DATA_YEAR <= 2016 ~ 'Barack Obama',
                                       DATA_YEAR >= 2017 ~ 'Donald Trump'))) %>% 
  select(president) %>%
  group_by(president) %>%
  summarize(count = sum(n()))%>%
  mutate(tenure = case_when(president == 'George Bush(1st term)'~ 2,
                                       president =='Bill Clinton'~ 8,
                                       president =='George Bush(2nd term)'~ 8,
                                       president =='Barack Obama'~8,
                                       president =='Donald Trump'~ 4)) %>%
  mutate(rate = count/tenure)

rate_plot = ggplot(president_ds,aes(x=fct_rev(fct_reorder(president,rate,.desc = TRUE)),y=rate)) +
  geom_bar(stat = "identity",fill = "cornflowerblue") +
  ggtitle("Hate Crime rate per Administration") +
  coord_flip() +
  ylab("Hate Crime Rate (per year)") + xlab("President")+
#  theme(panel.grid.major.x = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+ theme_economist() +
  geom_text(aes(label=rate), vjust=-0.3, size=3.5, color="black")

rate_plot
```

At the same time, we can also draw a trend chart of the changing relationship of prejudice over time during the tenure of different presidents. We found an increase in all kinds of prejudice during the first George Bush administration. During the period of Bill Clinton, the general trend of anti Black or African american was up, while the general trend of anti White was down. All kinds of prejudice in the second George Bush years were in a lull (neither rising nor falling). The overall trend of anti-Black or African american decline under Barack Obama administration is different from any other period. While anti-Black or African american trends have risen sharply under Donald Trump years, other kinds of prejudice have crept up.

```{r, fig.height = 10, fig.width = 24, fig.align = "center"}
bias_top_5 = bias %>% head(5)
president_prejudice <- hate_plusyr %>% 
mutate(president = as.factor(case_when(DATA_YEAR < 1993 ~ '1George Bush(1st term)',
                                       DATA_YEAR >= 1993 & DATA_YEAR <= 2000 ~ '2Bill Clinton',
                                       DATA_YEAR >= 2001 & DATA_YEAR <= 2008 ~ '3George Bush(2nd term)',
                                       DATA_YEAR >= 2009 & DATA_YEAR <= 2016 ~ '4Barack Obama',
                                       DATA_YEAR >= 2017 ~ '5Donald Trump')))%>% 
  filter(BIAS_DESC %in% bias_top_5$BIAS_DESC)%>%
  select(DATA_YEAR,president,BIAS_DESC) 

predident_names <- c(
                    `1George Bush(1st term)` = "GB
1st",
                    `2Bill Clinton` = "Bill Clinton",
                    `3George Bush(2nd term)` = "George Bush(2nd term)",
                    `4Barack Obama` = "Barack Obama",
                    `5Donald Trump` = "Donald 
Trump"
                    )

president_prejudice_plot = ggplot(president_prejudice, aes(DATA_YEAR, color = BIAS_DESC))+
 geom_line(stat = 'count',lwd = 3)+
 facet_grid(.~president, scales = 'free', space = 'free',labeller = as_labeller(predident_names))+
 scale_x_continuous(breaks = seq(1991, 2020, 1))+ labs(title="Relationship Between Prejudice over Time During Presidency", x = "Year", y = "Hate Crime Cases")+  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5))+ labs(colour = "Prejudice") +theme(legend.key.size = unit(x = 2, units = 'line'),
       legend.text = element_text(size =20),
       legend.title = element_text(size = 25),title = element_text(size = 30)) +theme(strip.text.x = element_text(size = 25))+
theme(legend.position = "top",axis.text.x=element_text(size=15),axis.text.y=element_text(size=15))

president_prejudice_plot
```



## Which groups of people have the highest hate crime cases?

Note here we ignore the unknown case since it tells nothing. Through this pie chart, we can find that 70.7 percent of all reported hate crime cases were committed by White people, which is nearly 45% higher than the black or African American group that ranks second. We can find that Asians are less likely to be offenders.

```{r, fig.align = "center"}
offender_df = hate_clean %>%select(OFFENDER_RACE) %>%
 group_by(OFFENDER_RACE) %>%
 summarize(count = sum(n())) %>%
 arrange(desc(count))
offender_df = offender_df[2:7,]
colors <- c('rgb(211,94,96)', 'rgb(128,133,133)', 'rgb(144,103,167)', 'rgb(171,104,87)', 'rgb(114,147,203)','rgb(46,8,8)','rgb(135,20,93)')

fig <- plot_ly(offender_df, labels = ~OFFENDER_RACE, values = ~count, type = 'pie',
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',
        marker = list(colors = colors,
                      line = list(color = '#FFFFFF', width = 1)),
                      #The 'pull' attribute can also be used to create space between the sectors
        showlegend = TRUE)
fig <- fig %>% layout(title = 'Number of Offenders by Race (Percentage) (1991-2020)',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```




   
## What types of hate crimes (Offense Type) were most common in the last 30 years?

This image shows all types of offense type. We found that "Destruction/Damage/Vandalism of Property", "Intimidation", "Simple Assault" , "Aggravated Assault" are the most popular. And the total number of incidents is far ahead of the latter category. We will use these four key offense type to draw the relationship between Offense type and Prejudice.

```{r, fig.height = 10, fig.width = 8}
offense_type <- hate_clean %>% select(OFFENSE_NAME) %>%
 group_by(OFFENSE_NAME) %>%
 summarize(count = sum(n()))%>%
 arrange(desc(count))

ggplot(offense_type, aes(x = count, y = fct_reorder(OFFENSE_NAME, count))) +
  geom_point(color = "blue") +
  labs(title="Total Hate Crimes per Offense Type (1991-2020)", x = "Hate Crime Cases", y = "Offense Type")+  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5))

```

We took the top four offense types in the figure above and the top seven prejudice types together to draw this grouped bar chart. This chart shows the total amount of hate crimes that were motivated by each type of prejudice for each type of hate crime. The majority of each type of hate crime is motivated by an Anti-Black or African American prejudice, with "Intimidation" and "Destruction/Damage/Vandalism of Property" being the most common type of hate crimes motivated by this prejudice. "Anti-Black or African American" prejudice has the highest percentage in every offense type. We found that the percentage of Anti-Jewish prejudice in the "Destruction/Damage/Vandalism of Property" offense type is particularly high. The LGBT community also makes up a significant portion of every offense type. 


```{r}
offense_type <- hate_clean %>% select(OFFENSE_NAME) %>%
 group_by(OFFENSE_NAME) %>%
 summarize(count = sum(n()))%>%
 arrange(desc(count))

offensetype_top_4 = offense_type %>% head(4)
offensetype_df = hate_clean %>% filter(OFFENSE_NAME %in% offensetype_top_4$OFFENSE_NAME) %>%
  filter(BIAS_DESC %in% bias_top_7$BIAS_DESC)  %>% left_join(offensetype_top_4)



fig <- offensetype_df %>% count(OFFENSE_NAME, BIAS_DESC) %>% left_join(offensetype_top_4)
fig <- fig %>% plot_ly(x = ~fct_rev(fct_reorder(OFFENSE_NAME,count,.desc = FALSE)), y = ~n, color = ~BIAS_DESC)%>%layout(title = 'Total Prejudice per Offense Type (1991-2020)', plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Offense Type'), 
         yaxis = list(title = 'Hate Crime Cases'), legend = list(title=list(text='<b> Type of Prejudice </b>')))

fig

```

This mosaic graph shows the relationship between Offender Race and Offense Type. Note this results ignore the unknown Unknown Offender Race. We found that Whites made up the largest portion of the top 4 offense types, followed by Blacks or African Americans.

```{r, fig.height = 15, fig.width = 18, fig.align = "center"}
library(ggmosaic)
offensetype_df_1  <- offensetype_df[offensetype_df$OFFENDER_RACE != "Unknown", ]
offensetype_df_2 = count(offensetype_df_1, OFFENSE_NAME, OFFENDER_RACE)


offensetype_df_2_mosaic = ggplot(data = offensetype_df_2) +
  geom_mosaic(aes(weight = n, x = product(OFFENSE_NAME), fill = OFFENDER_RACE))+ theme_grey(20)+ 
  theme(axis.text.x = element_text( 
    angle = 60)
  )+ggtitle("Offender Race vs Offense Type (1991-2020)") +xlab("Offense Type")+ylab("Offender Race")+guides(fill=guide_legend(title="Offender Race"))

offensetype_df_2_mosaic

```

## The relationship between total hate crime and month.

By observing the bar chart below, we can find that September and October are the periods of high incidence of hate crimes, while December is the period of low incidence of hate crimes

```{r}
hate_plus_month = hate_clean %>% mutate(Month = month(INCIDENT_DATE))
hate_plus_month_1 = hate_plus_month[!duplicated(hate_plus_month[,c('INCIDENT_ID','Month')]),]
month <- hate_plus_month_1 %>% select(Month) %>%
 group_by(Month) %>%
 summarize(count = sum(n()))
month = month %>% mutate(month_chr = as.factor(case_when(Month == 1 ~ 'Jan',
                                       Month == 2 ~ 'Feb',
                                       Month == 3 ~ 'Mar',
                                       Month == 4 ~ 'Apr',
                                       Month == 5 ~ 'May',
                                       Month == 6 ~ 'Jun',
                                       Month == 7 ~ 'Jul',
                                       Month == 8 ~ 'Aug',
                                       Month == 9 ~ 'Sep',
                                       Month == 10 ~ 'Oct',
                                       Month == 11 ~ 'Nov',
                                       Month == 12 ~ 'Dec')))

fig <- plot_ly(month, x = ~fct_rev(fct_reorder(month_chr,Month,.desc = TRUE)), y = ~count, type = 'bar',
             marker = list(color = 'rgb(158,202,225)',
                           line = list(color = 'rgb(8,48,107)', width = 1.5)))
fig <- fig %>% layout(title = "Hate Crime Totals per Month (1991-2020)",
         xaxis = list(title = "Month"),
         yaxis = list(title = "Hate Crime Cases"))%>% 
  layout(annotations = list(x = "Sep", y = 20738, text = "Max", showarrow = T)) %>%
  layout(annotations = list(x = "Dec", y = 14245, text = "Min", showarrow = T))

fig

```





## The relationship between hate crime cases and region. 

We found that the West region has the highest number of reported Hate Crime cases, and U.S. Territories has the least number. Here U.S. Territories stands for Guam which makes sense since it has small number of population. The Midwest has the lowest number of hate crimes in the continental United States. 

```{r}

hate_region_1 = hate_clean[!duplicated(hate_clean[,c('INCIDENT_ID','REGION_NAME')]),]
region <- hate_region_1 %>% select(REGION_NAME) %>%
 group_by(REGION_NAME) %>%
 summarize(count = sum(n()))

region_2 = region%>%mutate(Region_name = factor(REGION_NAME,levels=c("Other","U.S. Territories","Midwest","South","Northeast","West")))

fig <- plot_ly(region_2, x = ~count, y = ~Region_name, type = 'bar',orientation = 'h',
             marker = list(color = 'rgb(158,202,225)',
                           line = list(color = 'rgb(8,48,107)', width = 1.5)))
fig <- fig %>% layout(title = "Total number of hate crimes in the United States by Region (1991-2020)",
         xaxis = list(title = "Hate Crime Cases"),
         yaxis = list(title = "Region"))

fig

```













