# Results

## Trends of hate crime: Overall rise, but there were also periods of Depression

The following shows the trend chart of yearly reported hate crime cases reported. As can be seen from the figure, hate crime cases increased most sharply in 1991-1992, 1994-1995, 2000-2001, and 2019-2020. Conversely, the decline was the most dramatic from 1993-1994, 2001-2002, 2008-2009, and 2012-2013.

From 2001 to 2014, the total number of hate crime cases showed a general trend of decline. However, from 2014 to 2020, the total number of hate crime cases in the United States increased dramatically.

In the whole chart, 1991 had the lowest number of hate crime cases. On the other hand, 2020 has the highest total number of hate crime cases. 

We know the year has something to do with how long a U.S. president has been in office. Based on this picture, we will answer the question of which presidential administration had the highest or lowest hate crime incidents in the third subsection. 

```{r}
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggthemes)
library(plotly)
hate_plusyr = hate_clean %>% mutate(DATA_YEAR = year(INCIDENT_DATE))
hate_plusyr_1 = hate_plusyr[!duplicated(hate_plusyr[,c('INCIDENT_ID','DATA_YEAR')]),]
year <- hate_plusyr_1 %>% select(DATA_YEAR) %>%
 group_by(DATA_YEAR) %>%
 summarize(count = sum(n()))

trends_plot<- plot_ly(year, x = year$DATA_YEAR, y = year$count)
peak <- dplyr::filter(year, count == max(count))
troughs <- dplyr::filter(year, count == min(count))
second_peak = dplyr::filter(year, DATA_YEAR == 2001)
second_troughs = dplyr::filter(year, DATA_YEAR == 2014)
trends_plot %>%
  add_lines(x = year$DATA_YEAR, y = year$count) %>%
  layout(title = "Total Hate Crime Line Chart for 1991-2020",
         showlegend = FALSE,xaxis = list(title = 'Year'), 
         yaxis = list(title = 'Total case for reported Hate Crimes'))%>%
  layout(annotations = list(x = peak$DATA_YEAR, y = peak$count, text = "First Peak", showarrow = T)) %>%
  layout(annotations = list(x = troughs$DATA_YEAR, y = troughs$count, text = "First Trough", showarrow = T)) %>%
  layout(annotations = list(x = second_peak$DATA_YEAR, y = second_peak$count, text = "Second Peak", showarrow = T))%>%
  layout(annotations = list(x = second_troughs$DATA_YEAR, y = second_troughs$count, text = "Second Trough", showarrow = T))  
```

## Black or African American people were the most discriminated against

The plot shows the number of total hate crimes for each prejudice. As can be seen from the Cleveland dot chart, `Black or African Americans` are the groups most attacked by hate crime(almost 70,000 reported incidents). `Asians`, however, were the seventh most attacked group(over 7,000 reported incidents). Please note that our graph here only shows the total number of incidents involving people who have been attacked over the last 20 years, as opposed to the most vulnerable group. We will take the graph results to analyze the relationship between the total number of people attacked in different categories under different presidents. 

```{r, fig.height = 10, fig.width = 8, fig.align = "center"}
bias <- hate_clean %>% select(BIAS_DESC) %>%
 group_by(BIAS_DESC) %>%
 summarize(count = sum(n()))%>%
 arrange(desc(count))

ggplot(bias, aes(x = count, y = fct_reorder(BIAS_DESC, count))) +
  geom_point(color = "blue") +
  labs(title="Cleveland Dot Plot of Total Hate Crimes per Prejudice", x = "Hate Crime cases", y = "Prejudice")+  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5))
```

## Whites discriminated against other races the most, while Blacks preferred to discriminate against Whites

The following graph shows the relationship between the attacker's and the victim's race. Note we will ignore the `Unknown` offender race in this case. We found that when `Black or African American` people were attacked, `White` people attacked them the most, by a significant margin. We also found that `White` people were the most aggressive toward all groups except toward `Jewish`. Most `Anti-White` hate crimes were committed by `Black` and `White` people. 

```{r, fig.height = 10, fig.width = 18, fig.align = "center"}
bias_top_7 = bias %>% head(7)
attacker_attacked_df = hate_clean %>% filter(BIAS_DESC %in% bias_top_7$BIAS_DESC) %>% left_join(bias_top_7)

attacker_attacked_plot = ggplot(attacker_attacked_df, aes(x=fct_rev(fct_reorder(BIAS_DESC,count,.desc = TRUE)), fill = OFFENDER_RACE)) +
  geom_bar(position = "dodge") +
  ggtitle("Offender Race per Prejudice (1991-2020)") +
  coord_flip() +
  ylab("Hate Crime cases") + xlab("Prejudice")+
  theme(plot.title = element_text(hjust = 0.5)) +theme(legend.key.size = unit(x = 2, units = 'line'),
       legend.text = element_text(size =15),
       legend.title = element_text(size = 18),title = element_text(size = 30)) +theme(strip.text.x = element_text(size = 25))+
theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=18))+theme(legend.position = c(0.8, 0.3))+ theme(legend.background = element_rect(fill="lightblue",
                                  linetype="solid", 
                                  colour ="darkblue")) +guides(fill=guide_legend(title="Offender Race"))

attacker_attacked_plot
```

## The frequency of Hate Crimes is the highest under Donald Trump, but it is not obvious

Because each U.S. president has a different number of years in office, comparing total hate crime cases here would be unfair. So we divided the number of hate crime cases during each president's tenure by the number of years in office to get the hate crime rate for each president. We found that `Donald Trump` had the highest hate crime rate and `George H. W. Bush` had the lowest. But the difference between the three presidents with the highest rates is not obvious.

```{r}
president_ds <- hate_plusyr %>% 
mutate(president = as.factor(case_when(DATA_YEAR < 1993 ~ 'George H. W. Bush',
                                       DATA_YEAR >= 1993 & DATA_YEAR <= 2000 ~ 'Bill Clinton',
                                       DATA_YEAR >= 2001 & DATA_YEAR <= 2008 ~ 'George W. Bush',
                                       DATA_YEAR >= 2009 & DATA_YEAR <= 2016 ~ 'Barack Obama',
                                       DATA_YEAR >= 2017 ~ 'Donald Trump'))) %>% 
  select(president) %>%
  group_by(president) %>%
  summarize(count = sum(n()))%>%
  mutate(tenure = case_when(president == 'George H. W. Bush'~ 2,
                                       president =='Bill Clinton'~ 8,
                                       president =='George W. Bush'~ 8,
                                       president =='Barack Obama'~8,
                                       president =='Donald Trump'~ 4)) %>%
  mutate(rate = count/tenure)

rate_plot = ggplot(president_ds,aes(x=fct_rev(fct_reorder(president,rate,.desc = TRUE)),y=rate)) +
  geom_bar(stat = "identity",fill = "cornflowerblue") +
  ggtitle("Hate Crime rate per Administration") +
  coord_flip() +
  ylab("Hate Crime Rate (per year)") + xlab("President")+
#  theme(panel.grid.major.x = element_blank())+
  theme(plot.title = element_text(hjust = 0.5))+ theme_economist() +
  geom_text(aes(label=rate), vjust=-0.3, size=3.5, color="black")

rate_plot
```

At the same time, we can also draw a trend chart of the changing relationship of prejudice overtime during the tenure of different presidents. We found an increase in all kinds of prejudice during the `George H. W. Bush` administration. During the `Bill Clinton` period, the general trend of `Anti Black or African American` was up, while the general trend of `Anti White` was down. All kinds of prejudice in the `George W. Bush` years were in a lull (neither rising nor falling). The overall trend of `Anti-Black or African American` decline under the Barack Obama administration differs from any other period. While `Anti-Black or African American` trends have risen sharply under `Donald Trump` years, different kinds of prejudice have crept up.

```{r, fig.height = 10, fig.width = 24, fig.align = "center"}
bias_top_5 = bias %>% head(5)
president_prejudice <- hate_plusyr %>% 
mutate(president = as.factor(case_when(DATA_YEAR < 1993 ~ '1George H. W. Bush',
                                       DATA_YEAR >= 1993 & DATA_YEAR <= 2000 ~ '2Bill Clinton',
                                       DATA_YEAR >= 2001 & DATA_YEAR <= 2008 ~ '3George W. Bush',
                                       DATA_YEAR >= 2009 & DATA_YEAR <= 2016 ~ '4Barack Obama',
                                       DATA_YEAR >= 2017 ~ '5Donald Trump')))%>% 
  filter(BIAS_DESC %in% bias_top_5$BIAS_DESC)%>%
  select(DATA_YEAR,president,BIAS_DESC) 

predident_names <- c(
                    `1George H. W. Bush` = "GHWB",
                    `2Bill Clinton` = "Bill Clinton",
                    `3George W. Bush` = "George W. Bush",
                    `4Barack Obama` = "Barack Obama",
                    `5Donald Trump` = "Donald Trump"
                    )

president_prejudice_plot = ggplot(president_prejudice, aes(DATA_YEAR, color = BIAS_DESC))+
 geom_line(stat = 'count',lwd = 3)+
 facet_grid(.~president, scales = 'free', space = 'free',labeller = as_labeller(predident_names))+
 scale_x_continuous(breaks = seq(1991, 2020, 1))+ labs(title="Relationship Between Prejudice over Time During Presidency", x = "Year", y = "Hate Crime Cases")+  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5))+ labs(colour = "Prejudice") +theme(legend.key.size = unit(x = 2, units = 'line'),
       legend.text = element_text(size =20),
       legend.title = element_text(size = 25),title = element_text(size = 30)) +theme(strip.text.x = element_text(size = 25))+
theme(legend.position = "top",axis.text.x=element_text(size=15),axis.text.y=element_text(size=15))

president_prejudice_plot
```

## White people have the most recorded Hate Crimes

Note here we ignore the unknown case since it tells nothing. Through this bar chart, we can find that most of the reported hate crime cases were committed by `White` people, which is a lot higher than the `Black or African American` group that ranks second. We can find that `Asians` are less likely to be offenders.

```{r, fig.align = "center"}
offender_df = hate_clean %>%select(OFFENDER_RACE) %>%
 group_by(OFFENDER_RACE) %>%
 summarize(count = sum(n())) %>%
 arrange(desc(count))
offender_df = offender_df[2:7,]
colors <- c('rgb(211,94,96)', 'rgb(128,133,133)', 'rgb(144,103,167)', 'rgb(171,104,87)', 'rgb(114,147,203)','rgb(46,8,8)','rgb(135,20,93)')

fig <- plot_ly(offender_df, x = ~fct_rev(fct_reorder(OFFENDER_RACE,count,.desc = FALSE)), y = ~count, type = 'bar',
        marker = list(color = colors, width = 1.5))

fig <- fig %>% layout(title = "Number of Offenders by Race (1991-2020)",
         xaxis = list(title = "Offenders Race"),
         yaxis = list(title = "Hate Crime Cases"))%>% 
  layout(annotations = list(x = "White", y = 90720, text = "Most", showarrow = T)) %>%
  layout(annotations = list(x = "Native Hawaiian or Other Pacific Islander", y = 85, text = "Least", showarrow = T))

fig
```
   
## Top Offense Types of Hate Crimes

This image shows all types of offense types. We found that `Destruction/Damage/Vandalism of Property`, `Intimidation`, `Simple Assault` , and `Aggravated Assault` are the most popular. And the total number of incidents is far ahead of the latter category. Therefore, we will use these four key offense types to draw the relationship between Offense type and Prejudice.

```{r, fig.height = 10, fig.width = 8}
offense_type <- hate_clean %>% select(OFFENSE_NAME) %>%
 group_by(OFFENSE_NAME) %>%
 summarize(count = sum(n()))%>%
 arrange(desc(count))

ggplot(offense_type, aes(x = count, y = fct_reorder(OFFENSE_NAME, count))) +
  geom_point(color = "blue") +
  labs(title="Total Hate Crimes per Offense Type (1991-2020)", x = "Hate Crime Cases", y = "Offense Type")+  theme_linedraw() +
  theme(plot.title = element_text(hjust = 0.5))
```

We took the top four offense types in the figure above and the top seven prejudice types of drawing this grouped bar chart. This chart shows the total number of hate crimes motivated by each kind of prejudice for each type of hate crime. The majority of each type of hate crime is motivated by an `Anti-Black or African American` prejudice, with `Intimidation` and `Destruction/Damage/Vandalism of Property` being the most common type of hate crimes motivated by this prejudice. `Anti-Black or African American` prejudice has the highest percentage in every offense type. In addition, we found that the percentage of `Anti-Jewish` prejudice in the `Destruction/Damage/Vandalism of Property` offense type is particularly high. The `LGBT` community also makes up a significant portion of every offense type. 

```{r}
offense_type <- hate_clean %>% select(OFFENSE_NAME) %>%
 group_by(OFFENSE_NAME) %>%
 summarize(count = sum(n()))%>%
 arrange(desc(count))

offensetype_top_4 = offense_type %>% head(4)
offensetype_df = hate_clean %>% filter(OFFENSE_NAME %in% offensetype_top_4$OFFENSE_NAME) %>%
  filter(BIAS_DESC %in% bias_top_7$BIAS_DESC)  %>% left_join(offensetype_top_4)

fig <- offensetype_df %>% count(OFFENSE_NAME, BIAS_DESC) %>% left_join(offensetype_top_4)
fig <- fig %>% plot_ly(x = ~fct_rev(fct_reorder(OFFENSE_NAME,count,.desc = FALSE)), y = ~n, color = ~BIAS_DESC)%>%layout(title = 'Total Prejudice per Offense Type (1991-2020)', plot_bgcolor = "#e5ecf6", xaxis = list(title = 'Offense Type'), 
         yaxis = list(title = 'Hate Crime Cases'), legend = list(title=list(text='<b> Type of Prejudice </b>')))

fig
```

This mosaic graph shows the relationship between Offender Race and Offense Type. Note these results ignore the `Unknown` Offender Race. We found that `Whites` made up the most significant portion of the top 4 offense types, followed by `Blacks or African Americans`.

```{r, fig.height = 15, fig.width = 18, fig.align = "center"}
library(ggmosaic)
offensetype_df_1  <- offensetype_df[offensetype_df$OFFENDER_RACE != "Unknown", ]
offensetype_df_2 = count(offensetype_df_1, OFFENSE_NAME, OFFENDER_RACE)


offensetype_df_2_mosaic = ggplot(data = offensetype_df_2) +
  geom_mosaic(aes(weight = n, x = product(OFFENSE_NAME), fill = OFFENDER_RACE))+ theme_grey(20)+ 
  theme(axis.text.x = element_text( 
    angle = 60)
  )+ggtitle("Offender Race vs Offense Type (1991-2020)") +xlab("Offense Type")+ylab("Offender Race")+guides(fill=guide_legend(title="Offender Race"))

offensetype_df_2_mosaic
```

## The total number of Hate Crimes at the end or beginning of the year is relatively small

By observing the bar chart below, we can find that `September` and `October` are the periods of high incidence of hate crimes, while `December` is the period of low incidence of hate crimes

```{r}
hate_plus_month = hate_clean %>% mutate(Month = month(INCIDENT_DATE))
hate_plus_month_1 = hate_plus_month[!duplicated(hate_plus_month[,c('INCIDENT_ID','Month')]),]
month <- hate_plus_month_1 %>% select(Month) %>%
 group_by(Month) %>%
 summarize(count = sum(n()))
month = month %>% mutate(month_chr = as.factor(case_when(Month == 1 ~ 'Jan',
                                       Month == 2 ~ 'Feb',
                                       Month == 3 ~ 'Mar',
                                       Month == 4 ~ 'Apr',
                                       Month == 5 ~ 'May',
                                       Month == 6 ~ 'Jun',
                                       Month == 7 ~ 'Jul',
                                       Month == 8 ~ 'Aug',
                                       Month == 9 ~ 'Sep',
                                       Month == 10 ~ 'Oct',
                                       Month == 11 ~ 'Nov',
                                       Month == 12 ~ 'Dec')))

fig <- plot_ly(month, x = ~fct_rev(fct_reorder(month_chr,Month,.desc = TRUE)), y = ~count, type = 'bar',
             marker = list(color = 'rgb(158,202,225)',
                           line = list(color = 'rgb(8,48,107)', width = 1.5)))
fig <- fig %>% layout(title = "Hate Crime Totals per Month (1991-2020)",
         xaxis = list(title = "Month"),
         yaxis = list(title = "Hate Crime Cases"))%>% 
  layout(annotations = list(x = "Sep", y = 20738, text = "Max", showarrow = T)) %>%
  layout(annotations = list(x = "Dec", y = 14245, text = "Min", showarrow = T))

fig
```

## The West and Northeast regions have the highest Hate Crime records

We found that the `West region` has the highest reported Hate Crime cases, and `U.S. Territories` has the least number. Here `U.S. Territories` stands for Guam, which makes sense since it has a small population. The `Midwest` has the lowest number of hate crimes in the continental United States. The `Other` entry refers to incidents reported by law enforcement at the feral level, such as the FBI and military police. Due to their distinctness, we did not dig into them.

```{r}
hate_region_1 = hate_clean[!duplicated(hate_clean[,c('INCIDENT_ID','REGION_NAME')]),]
region <- hate_region_1 %>% select(REGION_NAME) %>%
 group_by(REGION_NAME) %>%
 summarize(count = sum(n()))

region_2 = region%>%mutate(Region_name = factor(REGION_NAME,levels=c("Other","U.S. Territories","Midwest","South","Northeast","West")))

fig <- plot_ly(region_2, x = ~count, y = ~Region_name, type = 'bar',orientation = 'h',
             marker = list(color = 'rgb(158,202,225)',
                           line = list(color = 'rgb(8,48,107)', width = 1.5)))
fig <- fig %>% layout(title = "Total number of hate crimes in the United States by Region (1991-2020)",
         xaxis = list(title = "Hate Crime Cases"),
         yaxis = list(title = "Region"))

fig
```
